---
title: "Predicting Airbnb prices"
author: "Nicolás Sandoval"
date: "28-07-2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(tidymodels)) install.packages("tidymodels", repos = "http://cran.us.r-project.org")
if(!require(textrecipes)) install.packages("textrecipes", repos = "http://cran.us.r-project.org")
#if(!require(textfeatures)) install.packages("textfeatures", repos = "http://cran.us.r-project.org")
if(!require(tidytext)) install.packages("tidytext", repos = "http://cran.us.r-project.org")
if(!require(lubridate)) install.packages("lubridate", repos = "http://cran.us.r-project.org")
if(!require(ggrepel)) install.packages("ggrepel", repos = "http://cran.us.r-project.org")
if(!require(ggmap)) install.packages("ggmap", repos = "http://cran.us.r-project.org")
if(!require(ggthemes)) install.packages("ggthemes", repos = "http://cran.us.r-project.org")
if(!require(magrittr)) install.packages("magrittr", repos = "http://cran.us.r-project.org")
if(!require(heatmaply)) install.packages("heatmaply", repos = "http://cran.us.r-project.org")

```

```{r load dataset}
dataset=read_csv("listings.csv")
dataset%<>%mutate(price=parse_number(price))
```

```{r}
set.seed(2807)
split=initial_split(dataset, prop = 0.85)
train_set=training(split)
test_set=testing(split)
```

```{r}
train_set%>%group_by(price)%>%
  summarize(count=n())
```
The data was scraped on July 4th, 2021.
The dataset contains 16724 observations of 74 variables, with a mixture of categorical and numerical information.
Some rows are not useful for this analysis, for example, the URL for the listing, the URL for the host's page, the URL for picture of the listing, etc. Some columns also contain no information, such as `bathrooms`
The goal of this project is to predict the price of each listing, based on the rest of the listing information.
The price column has character data ("$150.00"), but it's easy to convert to numeric via parse_number.
Prices go from 0 to 8000 USD per night.
```{r}
train_set%>%
  ggplot(aes(longitude, latitude, color=neighbourhood_cleansed))+
  geom_point(size=1)+scale_color_viridis_d()+theme(legend.position = "none") 
```

```{r}
train_set%>%ggplot(aes(price))+geom_density()
train_set%>%ggplot(aes(log(price+1)))+geom_density()
train_set%<>%mutate(price=log(price+1))
train_set%>%with(range(price))
train_mu=mean(train_set$price)


train_set%>%
  #mutate(price=parse_number(price))%>%
  group_by(latitude=round(latitude,2),
           longitude=round(longitude,2))%>%
  summarize(price=mean(price))%>%
  ggplot(aes(longitude, latitude, color=price))+
  geom_point()+scale_color_gradient2(high = "yellow", low = "blue", midpoint = train_mu)

price_tble=train_set%>%
 #mutate(price_num=parse_number(price))%>%
  select(price,price_num)
```

```{r}
train_head=train_set%>%head()
select_test=train_head%>%select(-contains("url"))
```

```{r}
train_set%>%group_by(neighbourhood_cleansed)%>%summarize(price=mean(price-train_mu))%>%
  ggplot(aes(price,reorder(neighbourhood_cleansed,price)))+geom_col()

train_set%>%group_by(room_type)%>%summarize(price=mean(price-train_mu))%>%
  ggplot(aes(price,reorder(room_type,price)))+geom_col()

train_set%>%filter(bedrooms<15)%>%
  ggplot(aes(bedrooms,price, group=bedrooms))+geom_boxplot()

train_set%>%filter(bedrooms<15)%>%
  ggplot(aes(bedrooms,price))+geom_smooth()

train_set%>%drop_na(bedrooms)%>%
  with(range(bedrooms))
train_set%>%
  mutate(bath_bed_ratio=parse_number(bathrooms_text)/bedrooms)%>%
  ggplot(aes(bath_bed_ratio,price))+geom_point()

#listing with shared baths have lower prices on average
train_set%>%mutate(bath_type=if_else(
  str_detect(bathrooms_text, "shared"), "Shared", "Private"))%>%
  group_by(bath_type)%>%summarize(price=mean(price))%>%
  ggplot(aes(bath_type,price-train_mu))+geom_col()

bath_parse=function(x){
  if_else(str_detect(x, regex("Half-bath", ignore_case = TRUE)), 0.5, parse_number(x)
  )
}

train_set%>%
  mutate(bathrooms=bath_parse(bathrooms_text), bath_bed_ratio=bathrooms/bedrooms)%>%
  ggplot(aes(bath_bed_ratio,price))+geom_point()

#listing with 0 bathrooms have low prices
train_set%>%
  mutate(bathrooms=bath_parse(bathrooms_text), no_bath=if_else(
    bathrooms==0, "no bath", "has baths"))%>%
  group_by(no_bath)%>%summarize(price=mean(price-train_mu))%>%
  ggplot(aes(no_bath,price))+geom_col()

train_set%>%
  group_by(host_id)%>%
  summarize(price=mean(price))%>%ggplot(aes(host_id,price))+
  geom_point()

train_set%>%
  group_by(property_type)%>%
  summarize(price=mean(price))%>%ggplot(aes(reorder(property_type,price), price))+
  geom_col()

train_set%>%ggplot(aes(minimum_nights,price))+geom_point()

train_set%>%ggplot(aes(number_of_reviews, price))+geom_point()

```

```{r}
bath_type=function(x){
  if_else(str_detect(x, "shared"), "Shared", "Private")
}

train_set%<>%mutate(shared_bath=as.factor(bath_type(bathrooms_text)),bathrooms=bath_parse(bathrooms_text), no_bath=if_else(
    bathrooms==0, "no bath", "has baths"))
train_set%>%group_by(no_bath)%>%summarize(count=n())
folds=vfold_cv(train_set)

```

```{r recipe}
location_rec=recipe(price~neighbourhood_cleansed+longitude+latitude,data=train_set)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)

loc_type_rec=recipe(price~property_type+room_type+neighbourhood_cleansed+longitude+latitude,data=train_set)%>%
  step_other(property_type)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)

type_rec=recipe(price~property_type+room_type,data=train_set)%>%
  step_other(property_type)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)

bath_location_rec=recipe(price~neighbourhood_cleansed+property_type+room_type+longitude+latitude+shared_bath+no_bath,data=train_set)%>%
  step_other(property_type)%>%
  step_impute_mode(shared_bath, no_bath)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)

comb_avail_rec=recipe(price~property_type+room_type+neighbourhood_cleansed+longitude+latitude+availability_30,data=train_set)%>%
  step_other(property_type)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)

accom_rec=recipe(price~property_type+room_type+neighbourhood_cleansed+longitude+latitude+availability_30+accomodates,data=train_set)%>%
  step_other(property_type)%>%
  step_dummy(all_nominal())%>%step_normalize(all_predictors())%>%step_normalize(price)


tidy(simple_rec)

```

```{r tuning}
met=metric_set(rmse)
grid_control=control_grid(extract=extract_model,save_pred = TRUE, save_workflow = TRUE)
```

```{r linear model}
lm_model = linear_reg(penalty = tune()) %>% set_engine("glmnet")
```
```{r workflow}
lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(accom_rec)

lm_tune=lm_wflow%>%  tune_grid(resamples=folds,metrics=met, grid=tibble(penalty=seq(0,0.01,0.0001)))
lm_tune%>%collect_metrics()%>%arrange(mean)

autoplot(lm_tune)

```


location information
0.975        
penalty 0

property and room type
0.890
penalty 0

previous 2 models together
0.861
penalty 0.001

combined models plus availability
 0.835
penalty 0

previous model plus accomodates
0.861
penalty 0.0001


```{r}
lm_fit=fit(lm_wflow, train_set)
```

```{r}
bt_model =
  boost_tree(mode="regression",
            mtry = tune(),
             trees = tune(),
             learn_rate = tune()
           #  ,learn_rate = tune(),tree_depth = tune(), min_n = tune(), loss_reduction = tune(), sample_size = tune()
             )%>%
  set_engine("xgboost")

bt_wflow = 
  workflow() %>% 
  add_model(bt_model) %>% 
  add_recipe(accom_rec)

bt_tune = bt_wflow %>%
  tune_grid(folds,
            metrics = met,
            control = grid_control,
            grid =crossing(mtry = c(7),
                            trees = seq(1000, 1500, 25), learn_rate=c(0.01)))

autoplot(bt_tune)

bt_tune %>%
  collect_metrics() %>%
  arrange(mean)

```



```{r}
xg_rec= recipe(price~., data=train_set)
  
  
  
  
library(heatmaply)



heatmaply_cor(d_cor)

```



```{r apply transformations to test set}
test_set%>%mutate(price=log(price+1))
```

